---
title: "Motor Trend MPG Analysis"
author: "Mikhail Romadanovskiy"
date: "25 Jul 2015"
output: rmarkdown::tufte_handout
---

# Executive summary
In this report, we will analyze `mtcars` data set and explore the relationship between a set of available variables and miles per gallon (MPG). The data was extracted from the 1974 **Motor Trend** US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).
There are 2 main questions to consider: what type of transmission is better for mpg, and how it can be discribed in numeric terms. 
The t-test shows that mpg performance difference between cars with automatic and manual transmission. And it is about 7 MPG more for cars with manual transmission than those with automatic transmission. 
We fit several linear regression models and select the one with highest Adjusted R-squared value. Given model shows that cars that are lighter in weight with a manual transmission and cars that are heavier in weight with an automatic transmission will have higher MPG values.


## Contex
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

```{r Preparation, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}
##Loading requirements and data.
packages <- c("data.table", "ggplot2", "dplyr", "xtable", "gclus", "car")
# sapply(packages, library, character.only=TRUE, quietly=TRUE)
sapply(packages, require, character.only=TRUE, quietly=TRUE)

knitr::opts_chunk$set(echo=FALSE, cache=TRUE, results = "asis")
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)

data(mtcars)
mtcars[1:3,]
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)
attach(mtcars)

result <- t.test(mpg~am)
pVal <- result$p.value
estVal <- result$estimate
```
Since p-Value is pVal the automatic and manual transmissions are from different populations. And the mean for MPG of manual transmitted cars is about result$estimate[[2]] - result$estimate[[1]] more than that of automatic transmitted cars.  

## Regression Model
Testing simple model.
```{r SimpleMpgModel}
simpleMpgModel <- lm(mpg~am, data=mtcars)
xtable(summary(simpleMpgModel)) # results="hide"
```
Simple model can explain **34%** of variance of MPG. Adjusted R-squared value indicates that we need to add other variables to the model to make better predictions.

## Finding proper set of explanation variables
Let's use all the availible data to make a regression model.
```{r FullRegressionModel}
fullModel <- lm(mpg~.,data=mtcars)
xtable(summary(fullModel))
```
Full regression model can explain **78%** of MPG variance, but none of the coeficients are significant.
We could find more effective set of variables to explain mpg.

```{r EffectiveModel, echo=TRUE, results="hide"}
stepModel <- step(fullModel, k=log(nrow(mtcars)))
xtable(summary(stepModel)) # results="hide"
mpgModel <- lm(mpg~wt+qsec+am, data=mtcars)
```

Linear model mpg ~ wt + qsec + am - 1 explains **83%** of MPG variance, but consists only significant at 5% s.l. coefficients. 

```{r CompareModels, echo=TRUE}
xtable(anova(mpgModel, simpleMpgModel, fullModel))
xtable(confint(mpgModel))

amIntWtModel<-lm(mpg ~ wt + qsec + am + wt:am, data=mtcars)
xtable(summary(amIntWtModel)) # results hidden
```  

The measure of how much an observation has effected the estimate of a regression coefficient, we get sum((abs(dfbetas(mpgModel)))>1)
Therefore, the above analyses meet all basic assumptions of linear regression and well answer the questions.  
\pagebreak

# Appendix: Figures  
```{r, fig.width=9, fig.height=8, fig.cap = "Selected Dimensions Pair Graph", fig.fullwidth = TRUE}
cpairs(mtcars %>% select(mpg, wt, am, qsec) , bg = c("red", "darkblue")[unclass(mtcars$am)],
       pch=22, gap=.2,
       upper.panel=panel.smooth,
       main="Pair Graph of Motor Trend Car Road Tests",
       cex=0.8, cex.lab=1, cex.axis=0.8, cex.main=1, cex.sub=1)
```  

```{r, fig.width=9, fig.height=8, fig.cap = "Full Pair Graph", fig.fullwidth = TRUE}
cpairs(mtcars, bg = c("red", "darkblue")[unclass(mtcars$am)],
       pch=22, gap=.2, 
       upper.panel=panel.smooth,
       main="Pair Graph of Motor Trend Car Road Tests",
       cex=0.8, cex.lab=1, cex.axis=0.8, cex.main=1, cex.sub=1)
```  

```{r, fig.width=9, fig.height=8, fig.cap = "Box Plot"}
boxplot(mpg ~ am, data=mtcars, 
        xlab="Transmission (0 = Automatic, 1 = Manual)", 
        ylab="MPG", col="darkblue",
        main="Boxplot of MPG vs. Transmission")
```  

```{r, fig.width=9, fig.height=8, fig.cap = "Residuals Analysis", fig.fullwidth = TRUE}
par(mfrow = c(2, 2))
plot(mpgModel)
```
