---
title: "Statistical Inference Project"
author: "MiRo"
date: "22 Jul 2015"
output: 
  pdf_document:
    toc: true
    highlight: zenburn
    
---

## Overview
The central limit theorem (CLT) states that, given certain conditions, the arithmetic mean
of a sufficiently large number of samples of independent random variables, each with a
well-defined expected value and well-defined variance, will be approximately normally
distributed, regardless of the underlying distribution. This paper demonstrates this
theorem by applying it to a population whose distribution is most definitely not Gaussian.
For this purpose, I will use the exponential distribution.

The exponential distribution can be simulated in R with `rexp(n, lambda)` where
`lambda` $\lambda$ is the rate parameter. The mean of exponential distribution is 
$1/\lambda$ and the standard deviation is also $1/\lambda$. For this simulation,
we set $\lambda=0.2$. In this simulation, we investigate the distribution of
averages of 40-100 numbers sampled from exponential distribution with $\lambda=0.2$.

## Main question
Illustrate via simulation and associated explanatory text the properties of the 
distribution of the mean of 40 exponentials.  You should
1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical
variance of the distribution.
3. Show that the distribution is approximately normal.

##Loading requirements and data.
```{r Preparation, cache=TRUE, warning=FALSE, message=FALSE, echo=TRUE, results='hide'}
packages <- c("base","data.table", "ggplot2", "dplyr", "scales", "vwr", "gridExtra")
sapply(packages, library, character.only=TRUE, quietly=TRUE)

set.seed(10001)
```

## Simulations
Exponential distributions are interesting in that frequencies of x approach infinity as x
approaches zero, and the frequencies approach zero as x approaches infinity. This is clearly
shown by drawing a histogram on a large sample of numbers from the population.

We are about to simulate 1000 tests with n=40 of exponential distribution. The mean of exponential distribution is $1/\lambda$ and the standard deviation also equals to $1/\lambda$.
```{r Simulations}
testsN <- 1000
n <- 40
lambda <- 0.2

meanTested <- 1/lambda
sdTested <- 1/lambda
alpha <- 0.05

sampleMeans <- c()
sampleSDs <- c()

for(i in 1:testsN){
  sample <- rexp(n, lambda)
  sampleMeans <- c(sampleMeans, mean(sample))
  sampleSDs <- c(sampleSDs, sd(sample))
}
```

## Testing and plotting data 

```{r MeanVSMean, warning=FALSE}
xfit <- seq(min(sampleMeans), max(sampleMeans), length=testsN)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))

gplot <- ggplot(as.data.frame(x=sampleMeans),aes(sampleMeans)) +
  geom_histogram(binwidth = 1/30, col="darkblue", bg="darkblue") +
  labs(main = "Distribution of averages of samples, lambda=0.2",
       ylab ="Count", xlab = "Observed means") +
  geom_vline(xintercept = 1/lambda, col="red") +
  geom_line(data=data.frame(xfit,yfit), aes_string(x="xfit", y="yfit"), width=0.1)

gplot

meanMeans <- mean(sampleMeans)
varMeans <- var(sampleMeans)

```

Theoretical mean is equal to `r meanTested`. Let's test our example mean (`r meanMeans`)
with t test.
Variation of sample means is equal to `r varMeans` where the theoretical variation of 
the distribution is $\sigma^2 / n = 1/(\lambda^2 n) = 1/(0.04 \times 40)$ = `r 1/(0.04 * 40)`.

Due to the central limit theorem, the averages of samples follow normal distribution. 
```{r}
qqnorm(sampleMeans)
qqline(sampleMeans)
```
