---
title: "DS-JHU-RM-W4"
author: "MiRo"
date: "25 Jul 2015"
output: html_document
---
# Regression Models Coursera
## Week 4
###Loading requirements and data.
```{r Preparation, cache=TRUE, warning=FALSE, message=FALSE}
packages <- c("data.table", "ggplot2", "dplyr", "scales")
sapply(packages, require, character.only=TRUE, quietly=TRUE)

options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')
```

## Quiz

### Question 1
Consider the space shuttle data ?shuttle in the MASS library.
Consider modeling the use of the autolander as the outcome 
(variable name use). Fit a logistic regression model with 
autolander (variable auto) use (labeled as "auto" 1) versus not
(0) as predicted by wind sign (variable wind). Give the estimated 
odds ratio for autolander use comparing head winds, labeled as 
"head" in the variable headwind (numerator) to tail winds (denominator).
```{r}
library(MASS)
dim(shuttle)
head(shuttle)
shuttle$newUse <- as.numeric(shuttle$use == "auto")
fit <- glm(newUse ~ as.factor(wind) - 1, data=shuttle, family="binomial")
odds <- exp(summary(fit)$coef)
odds[1] / odds[2] # 0.9686888
```

### Question 2
Consider the previous problem. Give the estimated odds ratio 
for autolander use comparing head winds (numerator) to tail 
winds (denominator) adjusting for wind strength from the variable magn.
```{r}
fit <- glm(newUse ~ as.factor(wind) + factor(magn) - 1, 
           family="binomial", data=shuttle)
summary(fit)$coef
exp(coef(fit))
odds <- exp(cbind(OddsRatio = coef(fit), confint(fit)))
odds[1] / odds[2] # 0.9684981
```

### Question 3
If you fit a logistic regression model to a binary variable, for 
example use of the autolander, then fit a logistic regression 
model for one minus the outcome (not using the autolander) what
happens to the coefficients?
```{r}
fit3 <- glm(I(1 - newUse) ~ as.factor(wind) - 1, 
            data=shuttle, family="binomial")
summary(fit3)$coef # -0.2513144 -0.2831263
summary(fit)$coef # 0.2513144 0.2831263
# The coefficients reverse their signs. 
```

### Question 4
Consider the insect spray data InsectSprays. Fit a Poisson model 
using spray as a factor level. Report the estimated relative rate
comapring spray A (numerator) to spray B (denominator).
```{r}
data("InsectSprays")
fit <- glm(count ~ spray - 1, data=InsectSprays, family="poisson")
summary(fit)$coef
rate <- exp(coef(fit))
rate[1] / rate[2] # 0.9456522
```


### Question 5
Consider a Poisson glm with an offset, t. So, for example, a model of
the form glm(count ~ x + offset(t), family = poisson) where x is a factor
variable comparing a treatment (1) to a control (0) and t is the natural
log of a monitoring time. What is impact of the coefficient for x if we
fit the model glm(count ~ x + offset(t2), family = poisson) where 
t2 <- log(10) + t? In other words, what happens to the coefficients if we 
change the units of the offset variable. (Note, adding log(10) on the log 
scale is multiplying by 10 on the original scale.)
```{r}
fit <- glm(count ~ as.factor(spray) + offset(log(count+1)), 
         family="poisson", data=InsectSprays)
fit2 <- glm(count ~ as.factor(spray) + offset(log(10)+log(count+1)), 
            family="poisson", data=InsectSprays)
summary(fit)$coef
summary(fit2)$coef  
# as.factor(spray)B  0.003512473
# The coefficient estimate is unchanged
```


### Question 6
Consider the data
x <- -5:5
y <- c(5.12, 3.93, 2.67, 1.87, 0.52, 0.08, 0.93, 2.05, 2.54, 3.87, 4.97)
Using a knot point at 0, fit a linear model that looks like a hockey stick
with two lines meeting at x=0. Include an intercept term, x and the knot 
point term. What is the estimated slope of the line after 0?
```{r}
x <- -5 : 5
y <- c(5.12, 3.93, 2.67, 1.87, 0.52, 0.08, 0.93, 2.05, 2.54, 3.87, 4.97)
knotPoint <- c(0)
spline <- sapply(knotPoint, function(knot) (x > knot) * (x - knot))
xMatrix <- cbind(1, x, spline)
fit <- lm(y ~ xMatrix - 1)
yhat <- predict(fit)
slope <- fit$coef[2] + fit$coef[3]
slope # 1.013
plot(x, y)
lines(x, yhat, col=2)
```