---
title: "Tooth Growth Analysis"
author: "Mikhail Romadanovskiy"
date: "20 Jul 2015"
output: rmarkdown::tufte_handout
---

# Overview
\newthought{The ToothGrowth data}[^dataset] consists of measurements of the mean length of the odontoblast cells harvested from 60 guinea pigs. These animals were fed a diet of 6 Vitamin C supplements. The Vitamin C was administered either in the form of Orange Juice (OJ) or chemically pure Vitamin C (VC). Each animal received the same daily dosage of Vitamin C
consistently.

This study performs an exploration of this data. The goal of the exploration is to see if we
can reasonably infer that either the method of administering vitamin C or the dosage has an
impact on the size of odontoblast cells.

```{r Preparation, cache=TRUE, warning=FALSE, eval=TRUE, message=FALSE, echo=FALSE, results='hide'}
##Loading requirements and data.
packages <- c("data.table", "ggplot2", "dplyr", "xtable","knitr")
# sapply(packages, library, character.only=TRUE, quietly=TRUE)
sapply(packages, require, character.only=TRUE, quietly=TRUE)
knitr::opts_chunk$set(echo=FALSE, cache=TRUE)
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)
```

## Loading and preparing data
For starters, let's run a summary of ToothGrowth to see what the contents of the dataset
are.
```{r LoadingData, warning=FALSE, results='asis'}
data("ToothGrowth")
xtable(summary(ToothGrowth),caption = "ToothGrowth Summary")
```
And true to the data description, we can see that there are 60 observations with two
supplements and three dosages.

From this quick analysis we can determine that there are two qualitative variables, the
supplement and dosage (supp and dose). And there is one quantitative variable; len.
Finally, let's compute some descriptive statistics on the quantitative variable.
```{r results='asis'}
x<-data.frame(mean=mean(ToothGrowth$len),var= var(ToothGrowth$len))
xtable(x)
```

## Data discovery
Plotting ToothGrowth data to have a quick overview.
```{r fig.width=9, fig.height=5, fig.cap = "Data overview", fig.fullwidth = TRUE}
gplot <- ggplot(ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
  geom_violin(scale = "area") +
  facet_grid(. ~ supp) +
  xlab("Dose in miligrams") +
  ylab("Tooth length") +
  guides(fill=guide_legend(title="Supplement type")) +
  scale_fill_brewer(palette="Set1") +
  theme_bw(base_size = 12)
gplot
```
Since intervals are not intersect, we could say that there are singificant difference 
in tooth lenth depends on dosage. We could even build a linear model to discribe this 
dependence, but not there is no need to do it for that assessment.
The figure above seems to indicate that the dosage is the most important factor in
determining the cell size. This assumption is based on the means for both OJ and VC being
26 microns when the dosage is 2.0.
\pagebreak

## Testing delivery methods
Now our aim is to test how tooth lenght depends on dosage and delivery method.
Is there any significant difference between these methods?
For every dosage t-test has been done there.
```{r fig.width=9, fig.height=5, fig.cap = "Data overview", fig.fullwidth = TRUE}
meansDT <- ToothGrowth  %>% group_by(supp, dose) %>% 
  summarise(means=mean(len)) %>% select(means, supp, dose)
gplot <- ggplot(ToothGrowth, aes(x=len, fill=supp)) +
  geom_density(kernel="gaussian") +
  facet_grid(. ~ dose) +
  geom_vline(aes(xintercept=meansDT$means), meansDT[meansDT$supp=="VC",], 
             colour="red", linetype = "longdash", binwidth=0.2) +
  geom_vline(aes(xintercept=meansDT$means), meansDT[meansDT$supp=="OJ",], 
             colour="blue", linetype = "longdash", binwidth=0.2) +
  xlab("Driver") +
  ylab("Freq") +
  guides(fill=guide_legend(title="Supplement type")) +
  scale_fill_brewer(palette="Set1") +
  theme_bw(base_size = 12)
gplot
```

```{r}
alpha <- 0.05
for (selDose in unique(ToothGrowth$dose)) {
  vcSet <- ToothGrowth %>% filter(supp=="VC", dose == selDose)
  ojSet <- ToothGrowth %>% filter(supp=="OJ", dose == selDose)
  t <- t.test(vcSet$len, ojSet$len, alternative = "two.sided", 
              conf.level = (1-alpha))
  print(paste("p-Value for dosage ", selDose, " equals to ", 
              round(t$p.value,2),".", sep=""))
}
```

# Assumptions and Conclusions
We can conclude from this analysis that some types of therapies produces better results in
growing cells in the teeth of guinea pigs. We have seen that with 95% certainty that
administering 2.0ml of vitamin C results in better cell growth than the average. In fact, we
see that in general it is better to administer the vitamins in orange juice as the vitamins can be given at 1.0ml or 2.0. 
It should be noted, that there is a fundamental assumption in this experiment. That is, the
vitamin C promotes odontoblast cell growth. There was no control group data, that is,
guinea pigs that were give no vitamin C. Therefore the conclusions here are drawn
specifically on orange juice vs pure vitamin; and at very specific levels.


[^dataset]: Source: C. I. Bliss (1952) The Statistics of Bioassay. Academic Press.
