---
title: "DS-JHU-RM-W3"
author: "MiRo"
date: "24 Jul 2015"
output: html_document
---

# Regression Models Coursera
## Week 3
###Loading requirements and data.
```{r Preparation, cache=TRUE, warning=FALSE, message=FALSE}
packages <- c("data.table", "ggplot2", "dplyr", "scales")
sapply(packages, require, character.only=TRUE, quietly=TRUE)

options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')
```

### Lecture
```{r}
library(datasets); data("swiss");require(graphics)
pairs(swiss, panel = panel.smooth, main="Swish data", col=3+(swiss$Catholic>50))

summary(lm(Fertility~.,data=swiss))
```

### Dummy variable
```{r}
data("InsectSprays")
summary(lm(count~spray, data=InsectSprays))$coef
summary(lm(count~
             I(1*(spray=="B")) + I(1*(spray=="C")) + 
             I(1*(spray=="D")) + I(1*(spray=="E")) +
             I(1*(spray=="F"))
           , data = InsectSprays))$coef

#Если хочется, чтобы коэф модели были представлены относительно 0, те являлись
#медианами, стоит использовать -1
summary(lm(count~spray-1, data=InsectSprays))$coef
```

### Simulate data
```{r}
n <- 100; t <- rep(c(0, 1), c(n/2, n/2)); x <- c(runif(n/2), runif(n/2)); 
beta0 <- 0; beta1 <- 2; tau <- 1; sigma <- .2
y <- beta0 + x * beta1 + t * tau + rnorm(n, sd = sigma)
plot(x, y, type = "n", frame = FALSE)
abline(lm(y ~ x), lwd = 2)
abline(h = mean(y[1 : (n/2)]), lwd = 3)
abline(h = mean(y[(n/2 + 1) : n]), lwd = 3)
fit <- lm(y ~ x + t)
abline(coef(fit)[1], coef(fit)[2], lwd = 3)
abline(coef(fit)[1] + coef(fit)[3], coef(fit)[2], lwd = 3)
points(x[1 : (n/2)], y[1 : (n/2)], pch = 21, col = "black", bg = "lightblue", cex = 2)
points(x[(n/2 + 1) : n], y[(n/2 + 1) : n], pch = 21, col = "black", bg = "salmon", cex = 2)
```



## Quiz 3.
### Question 1
Consider the mtcars data set. Fit a model with mpg as the outcome 
that includes number of cylinders as a factor variable and weight
as confounder. Give the adjusted estimate for the expected change
in mpg comparing 8 cylinders to 4.
```{r}
data(mtcars)
attach(mtcars)
cylF <- as.factor(cyl)
fit <- lm(mpg ~ cylF + wt, data=mtcars)
summary(fit)
x <- predict.lm(fit, data.frame(cyl = cylF, wt))
mean(x[cylF==4])-mean(x[cylF==8])
```

### Question 2
Consider the mtcars data set. Fit a model with mpg as the outcome
that includes number of cylinders as a factor variable and weight
as a possible confounding variable. Compare the effect of 8 versus
4 cylinders on mpg for the adjusted and unadjusted by weight models.
Here, adjusted means including the weight variable as a term in the
regression model and unadjusted means the model without weight 
included. What can be said about the effect comparing 8 and 4 cylinders
after looking at models with and without weight included?.
```{r}
fit2 <- lm(mpg ~ as.factor(cyl), data=mtcars)
summary(fit2) 
#$coef[3] # -11.56364
summary(fit)
#$coef[3] # -6.07086
# Holding weight constant, cylinder appears to have less of an 
# impact on mpg than if weight is disregarded.
```

### Question 3
Consider the mtcars data set. Fit a model with mpg as the outcome that 
considers number of cylinders as a factor variable and weight as confounder.
Now fit a second model with mpg as the outcome model that considers the 
interaction between number of cylinders (as a factor variable) and weight.
Give the P-value for the likelihood ratio test comparing the two models 
and suggest a model using 0.05 as a type I error rate significance benchmark.
```{r}
summary(fit)
fit3 <- lm(mpg ~ as.factor(cyl)*wt, data=mtcars)
# OR another way
# fit32 <- lm(mpg ~ factor(cyl) + wt + factor(cyl):wt, data=mtcars)
summary(fit3)
result <- anova(fit, fit3, test="Chi")
result$Pr # 0.1037502
# The P-value is larger than 0.05. So, according to our criterion, 
# we would fail to reject, which suggests that the interaction 
# terms may not be necessary.

```


### Question 4
Consider the mtcars data set. Fit a model with mpg as the outcome that 
includes number of cylinders as a factor variable and weight inlcuded in 
the model as
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
How is the wt coefficient interpretted?
```{r}
fit4 <- lm(mpg ~ I(wt * 0.5) + factor(cyl), data=mtcars)
summary(fit4)
# wt coef: The estimated expected change in MPG per one ton increase in 
# weight for a specific number of cylinders (4, 6, 8).

```


### Question 5
Consider the following data set
Give the hat diagonal for the most influential point
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
fit5 <- lm(y ~ x)
lm.influence(fit5)$hat[5] # 0.9945734
hatvalues(fit5)
```

### Question 6
Consider the following data set
Give the slope dfbeta for the point with the highest hat value.
```{r}
dfbetas(fit5) # -133.8226
```

### Question 7
Consider a regression relationship between Y and X with and 
without adjustment for a third variable Z. Which of the following
is true about comparing the regression coefficient between Y 
and X with and without adjustment for Z.
```{r}
# A: It is possible for the coefficient to reverse sign after 
# adjustment. For example, it can be strongly significant and 
# positive before adjustment and strongly significant and negative 
# after adjustment.
```
